Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-07-09 Delaneyetal2024.tex
small changes

While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
As a result, the sediment transport capacity response to individual extreme precipitation or melt events could be difficult to establish.
The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities.

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=\linewidth]{Fig7.pdf}
  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to water discharge variations under different channel conditions.
    a) Pipe flow with fixed channel size.
    b) R-channel described in Section \ref{sect:sub_mode}.
    c) Subaerial channel described in Section~\ref{sect:fluv}.
    d) Steady state R-channel described in Section~\ref{sect:scaling}.
    Note axes not to scale.
    Gyres on bottom panel represent shear stress on channel walls.
    $t_1$ ($t_2$) represents low (high) water discharge. 
  }
  \label{fig:chan_types}
\end{figure}

>>>>

While transport-limited states at glaciers are likely rare \citep[e.g.][]{alley1997}, abundant sediment can persist underneath glaciers \citep[e.g.][]{walter2014,stevens2022,delaney2022}.
As a result, in some cases sediment export can be hydrologically controlled and efficient \citep[e.g.][]{swift2005,andresen2024}.
At these glaciers, the great variability in subglacial sediment transport capacity may make it difficult to link sediment discharge to hydrology, especially when peak events occur \citep{cowan1988,delaney2018,lu2022}.
The results here suggest that different channel sizes with the same hydrology forcing can result in very different sediment transport capacities, making it challenging to establish the effects of individual extreme precipitation or melt events.

\begin{figure}[hbt!]
  \centering
  \includegraphics[width=\linewidth]{Fig7.pdf}
  \caption{Response of sediment transport capacity (orange line) and channel size (bottom row) to water discharge variations under different channel conditions.
    a) Pipe flow with fixed channel size described in Section~\ref{sect:scaling}.
    b) R-channel described in Section \ref{sect:sub_mode}.
    c) Subaerial channel described in Section~\ref{sect:fluv}.
    d) Steady state R-channel described in Section~\ref{sect:scaling}.
    Note axes not to scale.
    Gyres on bottom panel represent shear stress on channel walls.
    $t_1$ ($t_2$) represents low (high) water discharge. 
  }
  \label{fig:chan_types}
\end{figure}
